clusters

### 2015-07-13 cluster.rmd
heatmaps

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
all7c$Post.Code<- as.factor(all7c$Post.Code)

@@

x<- dist(clust %>% select(-mlsto), method = 'euclidean')
fit<- hclust(x, method = 'ward.D2')
plot(fit)

rect.hclust(fit, k=5, border='red')
#5 clusters looks pretty good

@@

#how many points in each cluster?
clust$groups<- cutree(fit,k=5)
table(clust$groups)

@@

```{r optimise clustnum,include=FALSE}
library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
library('vegan', lib = 'C:/Progra~1/R/R-3.2.0/library')

#try clustering a low number of numeric variables, trialling iffy ones to see which ones are good for clustering
clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      hrs.mlsto), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# client.totinv indicating about 4 clusters
# Num.days - good, indicating 18 custers
# hrs.mlsto - ok, heavy dip, good at about 17 clusters
# pc.pro - 5 groups good, still iffy
# mean.peeps - good at about 4 clusters
# Num.disc - ok, inidicating 12 groups
# timespan: not good for clustering
# no.users: not good for clustering
# hours.perday: no good for clustering

@@

k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto), 16)

@@

```{r CI, echo=FALSE}
library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')
# principal component analysis

@@

all7c$group.num<- as.integer(all7c$group.num)
p<- prcomp(~clust.inv + clust.bal+ clust.cost, data=all7c, centre=T)
# summary(p)
biplot(p, xlabs = as.character(all7c$group.num))
#putlers in-house biplot shows centroids
bpCent(p, clsAsgn = all7c$group.num, centroids=TRUE, xlabs = as.character(all7c$Business))

##try ggbiplot
library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
#visualise clusters
q<- ggbiplot(p, obs.scale = 1, var.scale = 1, 
              groups = all7c$group.num %>% as.factor, ellipse = TRUE,
             labels = all7c$Business,
              circle = TRUE)
q <- q + scale_color_discrete(name = '')
q <- q + theme(legend.direction = 'vertical', 
               legend.position = 'right')
q

@@

#confusion matrix of groups vs. categorical variables

@@

@@

>>>>

# library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
# library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
# library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
# library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')
# library('ggbiplot',lib='C:/Progra~1/R/R-3.2.1/library')
# library('RcmdrPlugin.BCA',lib='C:/Progra~1/R/R-3.2.1/library')

# library('RcmdrPlugin.BCA')
library(devtools)
library('ggbiplot')
library('fpc')
library('vegan')
library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')
library('party')
library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
# all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
all7c<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7c.csv')[,-1]
all7c$Post.Code<- as.factor(all7c$Post.Code)

@@

# x<- dist(clust %>% select(-mlsto), method = 'euclidean')
# fit<- hclust(x, method = 'ward.D2')
# plot(fit)
# 
# rect.hclust(fit, k=5, border='red')
#5 clusters looks pretty good

@@

#how many points in each cluster?
# clust$groups<- cutree(fit,k=5)
# table(clust$groups)

@@

```{r optimise clustnum,include=FALSE}

#try clustering a low number of numeric variables, trialling iffy ones to see which ones are good for clustering
# clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
#                                       hrs.mlsto), 2, 20, iter= 100)
# plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# client.totinv indicating about 4 clusters
# Num.days - good, indicating 18 custers
# hrs.mlsto - ok, heavy dip, good at about 17 clusters
# pc.pro - 5 groups good, still iffy
# mean.peeps - good at about 4 clusters
# Num.disc - ok, inidicating 12 groups
# timespan: not good for clustering
# no.users: not good for clustering
# hours.perday: no good for clustering

@@

k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto), 15)

@@

```{r CI, echo=FALSE}

# principal component analysis

@@

all7c$group.num<- as.integer(all7c$group.num)
p<- prcomp(~clust.inv + clust.bal+ clust.cost, data=all7c, centre=T)
# summary(p)
biplot(p, xlabs = as.character(all7c$group.num))
#putlers in-house biplot shows centroids
# bpCent(p, clsAsgn = all7c$group.num, centroids=TRUE, xlabs = as.character(all7c$Business))

##try ggbiplot

#visualise clusters
q<- ggbiplot(p, obs.scale = 1, var.scale = 1, 
              groups = all7c$group.num %>% as.factor, ellipse = TRUE,
             labels = all7c$group.num,
              circle = TRUE)
q <- q + scale_color_discrete(name = '')
q <- q + theme(legend.direction = 'vertical', 
               legend.position = 'right')
q

@@

#confusion matrix of groups vs. categorical variables
#reorder cluster tag levels from most amount and neg to best
all7c$group.num<- as.factor(all7c$group.num)
levs<- data.frame('clust' = rownames(clust.summary), 'bal' = clust.summary$balance.mlsto)
all7c$group.num <- factor(all7c$group.num, levels=levs[order(levs$bal),]$clust)

@@

#Biz.type
table(all7c$group.num, all7c$Biz.type)
#JD.Second table
table(all7c$group.num, all7c$JD.Second)
#combine JD.Second categories
all7c$JD.Second<- as.character(all7c$JD.Second)
#combine waste water and sewerage
all7c[all7c$JD.Second %in% 'sewerage',]$JD.Second<- 'waste water'
#combine re-use (water) and recycled water
all7c[all7c$JD.Second %in% 're-use (water)',]$JD.Second<- 'recycled water'
#combine water cycle management and water management and water strategy and flood studies
all7c[all7c$JD.Second %in% c('water cycle management', 'water strategy', 'flood studies'),]$JD.Second<-
        'water management'
#combine apartments and townhouses and community housing
all7c[all7c$JD.Second %in% c('townhouses','community housing','apartments'),]$JD.Second<- 'multi-dwelling'
#combine childcare and community
all7c[all7c$JD.Second %in% c('childcare', 'religious'),]$JD.Second<- 'community'
#combine commercial shop and dining
all7c[all7c$JD.Second %in% c('commercial shop','dining'),]$JD.Second<- 'dining/retail'
all7c$JD.Second<- as.factor(all7c$JD.Second)
all7c$JD.Second<- droplevels(all7c$JD.Second)

#Biz.type
table(all7c$group.num, all7c$pc.majpos)
```

@@

```{r heatmaps, echo=FALSE}
#function for heatmap

heatmap<- function(colour = 'deeppink3', clust = 'group.num', var = 'JD.Second', df=all7c, percentof_clust=FALSE){

        #order var levels by highest amount invoiced total
        a = ddply(df, c(var), summarise, summ = sum(inv.mlsto))
        df[,var] = factor(df[,var], levels= a[order(-a$summ),][,var])

        if(percentof_clust== FALSE){
                #table with percentages within each category
                tbl= round(prop.table(table(df[,clust],df[,var]),1)*100,2) %>% as.data.frame
                p = ggplot(tbl, aes(x=Var2, y=Var1)) +
                        geom_tile(aes(fill=Freq), colour = 'white') +
                        scale_fill_gradient(low='white', high=colour, name = 'percent') + 
                        labs(x='', y='', title= 'percentage of jobs in each cluster') +
                        theme(axis.ticks=element_blank(),
                              axis.text.x=element_text(angle=45,hjust=1))+
                        scale_x_discrete(expand = c(0, 0)) + scale_y_discrete(expand = c(0, 0)) 
        }
        else{
                #table with percentages within each cluster
                tbl= round(prop.table(table(df[,clust],df[,var]),2)*100,2) %>% as.data.frame
                p = ggplot(tbl, aes(x=Var2, y=Var1)) +
                        geom_tile(aes(fill=Freq), colour = 'white') +
                        scale_fill_gradient(low='white', high=colour, name = 'percent') + 
                        labs(x='Profit measure', y='', title= 'percentage of jobs in each X-axis category') +
                        theme(axis.ticks=element_blank(),
                              axis.text.x=element_text(angle=45,hjust=1))+
                        scale_x_discrete(expand = c(0, 0)) + scale_y_discrete(expand = c(0, 0))       
        }
        print(p)
}

heatmap(colour = 'deeppink3', clust = 'group.num', var = 'JD.Second', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'JD.Second', df=all7c, percentof_clust=TRUE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Business', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Business', df=all7c, percentof_clust=TRUE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Num.disc', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'majority.pos', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'majority.pos', df=all7c, percentof_clust=TRUE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Billing.Type', df=all7c, percentof_clust=TRUE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Discipline', df=all7c, percentof_clust=TRUE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'Discipline', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.client', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.client', df=all7c, percentof_clust=TRUE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.contact', df=all7c, percentof_clust=FALSE)
heatmap(colour = 'deeppink3', clust = 'group.num', var = 'code.contact', df=all7c, percentof_clust=TRUE)
